NanoMaLy

### 2015-11-12 paper/pldi16-submission/semantics.tex
compatible instead of refinement

We also define a refinement relation between types.

\begin{defn}
\label{defn:subtyping}
  A type $s$ is a \emph{refinement} of a type $t$, written \tsub{s}{t}, if
  $\exists \su.\ t = \subst{\su}{s}$.
\end{defn}

@@

\begin{lem}
\label{lem:refine-partial}
  For all $k$, \tsub{\ptype{k+1}{f}}{\ptype{k}{f}}.
\end{lem}

@@

  Let $s_k = \ptype{k}{f}$. Suppose \hastype{v}{\tnsub{t}{s_{k+1}}}, we
  will show that \stepi{k+1}{\eapp{f}{v}}{\emptysu}{\stuck}{\su}.

  \begin{description}
  \item[Case \tnsub{t}{s_k}:] 
    The inductive hypothesis applies.
  \item[Case \tsub{t}{s_k} but \tnsub{t}{s_{k+1}}:]
    By \autoref{cor:force} we must have called \forcesym at step $k+1$. A
    case analysis of the applicable rules shows that \forcesym cannot
    have succeeded.
  \end{description}
\end{proof}

@@

\begin{thm}
\label{thm:soundness}
  For any function $f$, if \steps{\eapp{f}{\ehole{}}}{\emptysu}{\stuck}{\su},
  then

$$
  \forall t. \exists \hastype{v}{t}. \steps{\eapp{f}{v}}{\emptysu}{\stuck}{\su}
$$
\end{thm}
\begin{proof}
  Let $\tfun{s}{\_} = \ptype{}{f}$.
  \begin{description}
  \item[Case \tsub{s}{t} or \tsub{t}{s}:] Our witness is already valid.
  \item[Case \tincompat{s}{t}:] We can apply Lemma~\ref{lem:k-stuck} and
    use \emph{any} \hastype{v}{t} as a witness for $f$.
  \end{description}
\end{proof}

>>>>

We also define a compatibility relation between types.

\begin{defn}
\label{defn:type-compat}
  A type $s$ is \emph{compatible} with a type $t$, written \tcompat{s}{t}, if
  $\exists \su.\ t = \subst{\su}{s} \lor s = \subst{\su}{t}$.
\end{defn}

@@

\begin{lem}
\label{lem:refine-partial}
  For all $k$, \tcompat{\ptype{k}{f}}{\ptype{k+1}{f}}.
\end{lem}

\begin{lem}
\label{lem:force-inst}
  For all $k$, if $\ptype{k}{f} \neq \ptype{k+1}{f}$, then \forcesym must
  have been called at step $k+1$.
\end{lem}
\begin{proof}
  By case analysis on the evaluation rules. If 
  $\ptype{k}{f} \neq \ptype{k+1}{f}$ then one of the holes in 
  $f$'s argument must have been instantiated with a concrete value at step $k+1$.
  An examination of the rules shows that only place this happens is 
  in the first case of \forcesym.
\end{proof}

@@

  Let $s_k = \ptype{k}{f}$. Suppose \hastype{v}{\tincompat{t}{s_{k+1}}}, we
  will show that \stepi{k+1}{\eapp{f}{v}}{\emptysu}{\stuck}{\su}.

  \begin{description}
  \item[Case \tincompat{t}{s_k}:] 
    The inductive hypothesis applies.
  \item[Case \tcompat{t}{s_k} but \tincompat{t}{s_{k+1}}:] 
    By Lemma~\ref{lem:force-inst} we must have called \forcesym at step
    $k+1$. A case analysis of the applicable rules shows that \forcesym
    cannot have succeeded.
  \end{description}
\end{proof}

@@

\begin{thm}
\label{thm:soundness}
  For any function $f$, if \steps{\eapp{f}{\ehole{}}}{\emptysu}{\stuck}{\su},
  then $\forall t. \exists \hastype{v}{t}. \steps{\eapp{f}{v}}{\emptysu}{\stuck}{\su}$.
\end{thm}
\begin{proof}
  Let $\tfun{s}{\_} = \ptype{}{f}$.
  \begin{description}
  \item[Case \tcompat{s}{t}:] Our witness is already valid.
  \item[Case \tincompat{s}{t}:] We can apply Lemma~\ref{lem:k-stuck} and
    use \emph{any} \hastype{v}{t} as a witness for $f$.
  \end{description}
\end{proof}
